An Introduction to Return Oriented Programming

### 2023-02-16 paper.tex
changes

\Verb+pop+ allows us to write arbitrary values into registers. For that we search for a \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value that we want to insert after the address of the \Verb+pop+ instruction. If we can not find a suitable gadget we can try to get creative and achieve the desired state another way. For example if we want to modify \Verb+ecx+ but do not have a \Verb+pop ecx+ instruction available we could achieve it with something like this: \bltInlineVerb{xor ecx, ecx ; pop eax ; xor ecx, eax}. Provided that we have these gadgets available.

@@

\Verb+mov+ allows us to write arbitrary values into memory. For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations. The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \bltInlineVerb{pop ecx ; pop eax ; mov dword ptr [eax], ecx}

@@

Arithmetic operations like \Verb+add+, \Verb+sub+, \Verb+inc+, \Verb+xor+, \Verb+or+, and can be useful to bring registers into our desired state. For that we search for the corresponding gadget with the required operands. For example \Verb+xor+ can be used to clear a register or copy its contents. It often occurs in the following forms: \bltInlineVerb{xor eax, eax} or \bltInlineVerb{xor eax, edx}. The first case clears the register since \Verb+xor+ computes a non-equivalence, formally $a \oplus a = 0$ and the second one copies the value of the 2nd operand into the 1st operand when the target register is \Verb+0x00+ since \Verb+0x00+ is the neutral element of the \Verb+xor+ operation, formally $a \oplus 0 = a$.

@@

\Verb+int+ stand for an interrupt, the interrupt \Verb+0x80+ causes a system call to be executed. System calls are kernelspace programs/operations that require higher privileges than what is available in a userspace program. Examples for system calls include io and \Verb+execve+ which allows to execute arbitary programs. In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call. One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+. This will be demonstrated in~\cref{sec:attack}

@@

for all of these regular expressions i was able to find at least a few suitable candidates. If there are no results the amount of possible characters before or after the gadget can be increased until results show up. It is however desirable to have gadgets with as few and noninterfering instructions as possible, if this is accomplished we can almost use the instructions we found like in assembly. Gadgets which do multiple things at once however can mess up the desired state and break the payload so it is important to thoroughly analyze the gadgets before using them.

\section{Theory}
\subsection{Stack}
The following graphic is an illustration of how the stack changes when injecting the payload. The buffer first has to be filled. In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}. It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s, because of that the buffer has to be filled with 16 Bytes instead of 8 Bytes, this can be turned off with the option \bltInlineVerb{-mpreferred-stack-boundary=2}. Though, then the payload only worked when filling the buffer with 24 Bytes.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.79\textwidth]{stackropoffsec.png}
  \caption{The stack when injecting the payload}
  \label{fig:mesh1}
\end{figure}
\subsection{ROP Chain}

@@

After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this its favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload. For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call. The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}

>>>>

\Verb+pop+ allows us to write arbitrary values into registers. For that we search for a \Verb+pop <reg>+ instruction inside our gadgets, in the payload we can then place the value that we want to insert after the address of the \Verb+pop+ instruction.~\cite{ropsla} If we can not find a suitable gadget we can try to get creative and achieve the desired state another way. For example if we want to modify \Verb+ecx+ but do not have a \Verb+pop ecx+ instruction available we could achieve it with something like this: \bltInlineVerb{xor ecx, ecx ; pop eax ; xor ecx, eax}. Provided that we have these gadgets available.

@@

\Verb+mov+ allows us to write arbitrary values into memory. For that we search for a \Verb+mov dword ptr [<reg1>], <reg2>+ instruction inside our gadgets, we can then, in combination with two pops write arbitrary values at arbitary memory locations.~\cite{ropsla} The following example writes the value in \Verb+ecx+ to where \Verb+eax+ points to: \bltInlineVerb{pop ecx ; pop eax ; mov dword ptr [eax], ecx}

@@

Arithmetic operations like \Verb+add+, \Verb+sub+, \Verb+inc+, \Verb+xor+, \Verb+or+, and can be useful to bring registers into our desired state.~\cite{ropsla} For that we search for the corresponding gadget with the required operands. For example \Verb+xor+ can be used to clear a register or copy its contents. It often occurs in the following forms: \bltInlineVerb{xor eax, eax} or \bltInlineVerb{xor eax, edx}. The first case clears the register since \Verb+xor+ computes a non-equivalence, formally $a \oplus a = 0$ and the second one copies the value of the 2nd operand into the 1st operand when the target register is \Verb+0x00+ since \Verb+0x00+ is the neutral element of the \Verb+xor+ operation, formally $a \oplus 0 = a$.

@@

\Verb+int+ stands for interrupt, the interrupt \bltInlineVerb{int 0x80} causes a system call to be executed. System calls are kernelspace programs/operations that require higher privileges than what is available in a userspace program. Examples for system calls include io and \Verb+execve+ which allows to execute arbitary programs. In combination with \Verb+pop+, \Verb+mov+ and other instructions we can specify the concrete system call.~\cite{ropsla} One of the most powerful system calls for blackhats is bash since it allows permanently implementing malware or gain insight into files, it can be called with the argument \Verb+/bin/sh+. This will be demonstrated in~\cref{sec:attack}

@@

for all of these regular expressions there were gadgets for the given program in~\cref{sec:attack}. If there are no results the amount of possible characters before or after the gadget can be increased until results show up. It is however desirable to have gadgets with as few and noninterfering instructions as possible, if this is accomplished we can almost use the instructions we found like in assembly. Gadgets which do multiple things at once however can mess up the desired state and break the payload so it is important to thoroughly analyze the gadgets before generating the payload.

\section{Theory}
\subsection{Stack}
The following graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload. The buffer first has to be filled. In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}. It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s, because of that the buffer has to be filled with 16 Bytes instead of 8 Bytes, this can be turned off with the option \bltInlineVerb{-mpreferred-stack-boundary=2}. Though, then the payload only worked when filling the buffer with 24 Bytes.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.79\textwidth]{stackropoffsec.png}
  \caption{The stack when injecting the payload}
  \label{fig:stack}
\end{figure}
\subsection{ROP Chain}

@@

After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this its favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload. For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call. The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}~\cite{proggen-rop}
